---
title: "STAT 302, Week 1"
subtitle: "Introduction and Getting Started with R"
author: "Bryan Martin"
date: "Jan 6, 2020"
output:
  xaringan::moon_reader:
    css: ["default", "metropolis", "metropolis-fonts", "my-theme.css"]
    lib_dir: libs
    nature:
      highlightStyle: tomorrow-night-bright
      highlightLines: true
      countIncrementalSlides: false
      titleSlideClass: ["center","top"]
---

```{r setup, include=FALSE, purl=FALSE}
options(htmltools.dir.version = FALSE)
knitr::opts_chunk$set(comment = "##")
```


# Outline

1. Course Overview
2. Introduction to R, RStudio, and RMarkdown
3. Getting started with data

.middler[**Goal:** Create a function Rmarkdown document (Short Lab 1)]

---

# Syllabus

.middler[.large[
[Link to syllabus](https://bryandmartin.github.io/STAT302/syllabus.html)
]]
---

# Expectations

What you should expect from me...
* your learning will be my priority
* you will be treated like an adult and with respect
* your feedback will be valued
* timely feedback on assignments
* understandable and well-paced lectures (tell me if they are not!)
* an attempt at making statistical computation as fun for you as it is for me!

---

# Expectations

What I will expect from you...
* regular attendance
* timely assignments that represent your best work
* a respectful and engaged classroom
* a desire and effort to learn challenging material

---

# Canvas Discussion

* Worth up to 5% extra credit on your final grade!
* Substantive and helpful questions and answers

--

.pull-left[
### Bad questions:

* How do you do problem 2?
* Here's my code and it's broken. How do I fix it?
]

.pull-right[
### Good questions:
* Here's a snippet of code I used for problem 2: 
<br/>`formatted code snippet`
<br/>It returned the following error:
<br/>`formatted error message`
<br/>Does anyone know why? I already tried...
* I don't understand the concept from Slide 18 today. Could anyone elaborate on why...
]

---

# Canvas Discussion

* Worth up to 5% extra credit on your final grade!
* Substantive and helpful questions and answers

.pull-left[
### Bad answers:
* Here's my solution
* Fool! You should already know the answer to this! Your trivial question is no match for my superior intellect!
]

.pull-right[
### Good answers:
* This error message occurs because your variable is a string instead of a numberic.
Have you tried checking...
* I think you have a bug in line 3 of the code you posted. You have more left parenthesis than right parenthesis so the line is incomplete.
]

---

# Why R?

R is a programming language designed for statistical analysis.

--

* open-source
* free
* large and active community of developers and users
* great analysis tools
* great visualization tools
--

* great user interface...

---

# Why RStudio?

RStudio is an integrated development environment (IDE) designed to make your life easier.

--

* Organizes scripts, files, plots, code console, ...
* Highlights syntax
* Helpful interactive graphical interface
* Will make an efficient, reproducible workflow *much* easier
--

* Rmarkdown integration...

---

# Why Rmarkdown?

* Interface between code, output, and writing
* Self-contained analyses
* Creates HTML, PDF, slides (like these!), webpages, ...
--

* Required for your labs!

---

# Operators

```{r}
# Addition
6 + 3
```

```{r}
# Subtraction
6 - 3
```

```{r}
# Multiplication
6 * 3
```

```{r}
# Division
6 / 3
```


---


# Comparison Operators

```{r}
# Greater than
6 > 3
```

```{r}
# Less than
6 < 3
```

```{r}
# Equal to
6 == 3
6 == 3 + 3
```


---

# Comparison Operators

```{r}
# Not equal to
6 != 3
```

```{r}
6 < 6
# Less than or equal to
6 <= 6
```


---

# Logical Operators

```{r, eval = FALSE}
# and
(6 < 3) & (1 < 3)
```


--


```{r, echo = FALSE}
# and
(6 < 3) & (1 < 3)
```


--

```{r, eval = FALSE}
# and
(2 < 3) & (1 < 3)
```


--


```{r, echo = FALSE}
# and
(2 < 3) & (1 < 3)
```


--

```{r, eval = FALSE}
# or
(6 < 3) | (1 < 3)
```


--


```{r, echo = FALSE}
# or
(6 < 3) | (1 < 3)
```


--

```{r, eval = FALSE}
# a bit harder...
(6 < 3) | (1 < 3) & (6 < 3)
```


--


```{r, echo = FALSE}
# a bit harder...
(6 < 3) | (1 < 3) & (6 < 3)
```


---

# Object Types

```{r}
class(7)
class("7")
is.numeric(7)
is.numeric("7")
```


---

# Object Types

```{r}
is.character(7)
is.character("7")
is.na(7)
is.na(0/0)
```


---

# Object Types

```{r, error = TRUE}
as.character(7)
as.numeric("7")
as.numeric("7") + 3 == 10
"7" + 3 == 10
```


---

# Assigning Variables

```{r}
x <- 7
x
x + 3
x == 7
as.character(x)
y <- 3
x + y
```


---

# Workspaces

```{r, error = TRUE}
# List all defined objects
ls()
# Remove an object
rm("x")
ls()
x
```


---

# Workspaces

```{r}
x <- 7
ls()
# Use with caution! This erases everything!
rm(list = ls())
ls()
```

