<!--- Begin styling code. --->
<style type="text/css">
/* Whole document: */
body{
  font-family: "Palatino Linotype", "Book Antiqua", Palatino, serif;
  font-size: 12pt;
}
h1.title {
  font-size: 38px;
  text-align: center;
}
h4.author {
  font-size: 18px;
  text-align: center;
}
h4.date {
  font-size: 18px;
  text-align: center;
}
</style>
<!--- End styling code. --->

---
title: "Lab 2"
author: "STAT 302"
date: "1/20/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

*If you collaborated with anyone, you must include "Collaborated with: FIRSTNAME LASTNAME" at the top of your lab!*

## Part 1. More on Functions (8 points)

**1a.** (2 points) The hard-threshold function is defined as 
$$f_\lambda(x) = \begin{cases} x & |x| \geq \lambda\\ 0 & |x| < \lambda \end{cases}$$
Write an R function that takes two parameters, numeric input `x` and a threshold `lambda`.
Your function should return the value of $f_\lambda(x)$ and work for vector input `x` of any length.

**1b.** (1 point) Set $\lambda = 4$, demonstrate your function on the vector `c(-5, -3, 0, 3, 5)`.

(*Hint: the output should be the vector `-5, 0, 0, 0, 5`*)

**1c.** (1 point) Set $\lambda = 2$, demonstrate your function on the vector `c(-7, -5, -3, 0, 3, 5, 7)`.

**2a.** (2 points) The soft-threshold function is defined as 
$$g_\lambda(x) = \begin{cases} sign(x)(|x| - \lambda) & |x| \geq \lambda\\ 0 & |x| < \lambda \end{cases}$$
Write an R function that takes two parameters, numeric input `x` and a threshold `lambda`.
Your function should return the value of $g_\lambda(x)$ and work for vector input `x` of any length.

**2b.** (1 point) Set $\lambda = 4$, demonstrate your function on the vector `c(-5, -3, 0, 3, 5)`.

(*Hint: the output should be the vector `-1, 0, 0, 0, 1`*)

**2c.** (1 point) Set $\lambda = 2$, demonstrate your function on the vector `c(-7, -5, -3, 0, 3, 5, 7)`.

## Part 2. Lists (4 points)

Many popular functions in R output lists in order to return multiple objects of different types and lengths. Here we will look at the function `lm`, which performs linear regression. 
Don't worry if you aren't familiar with linear regression, we will just be analyzing the output of the function.

First, run the following code to create an object of class `lm`.

```{r}
linearMod <- lm(dist ~ speed, data = cars) 
```

**3a.** (2 points) What are the names of the items in the list `linearMod`?

**3b.** (2 points) Use two different methods to store the `coefficients` within `linearMod` as a variable. One method should use the dollar sign `$`, the other should use the word `coefficients`. Print out your variables.

## Part 3. Data (8 points)

**4.** (2 points) Create a data frame that includes at least 3 observations and variables of at least 3 different types. Print out your data frame.

For problem 5, we will use an adapted version of the weather data from Tidyverse.

```{r}
library(kableExtra)
weather <- data.frame("station" = rep(c("A", "B", "C"), each = 4),
                      "element" = rep(c("temp_min", "temp_max"), 2),
                      "month1"  = c(11.4, 25.6, NA, NA, 17.7, 28.0,
                                    NA, NA, 20.0, 24.9, NA, NA),
                      "month2"  = c(NA, NA, 16.8, 28.7, NA, NA,
                                    11.1, 26.8, NA, NA, 14.7, 33.4))
kable_styling(kable(weather))
```

**5a.** (2 points) Identify and list the observations and the variables in this data set.

(*Hint: the variables are not `station`, `element`, `month1`, and `month2`. These are the column names in the untidy data, and need to be fixed!*)

**5b.** (4 points) Use `kable()` to present a tidied up version of this data. (Hint: My table has 4 variables, 6 observations, and no `NA` values. I only kept one of the variables unchanged.)

