<!--- Begin styling code. --->
<style type="text/css">
/* Whole document: */
body{
  font-family: "Palatino Linotype", "Book Antiqua", Palatino, serif;
  font-size: 12pt;
}
h1.title {
  font-size: 38px;
  text-align: center;
}
h4.author {
  font-size: 18px;
  text-align: center;
}
h4.date {
  font-size: 18px;
  text-align: center;
}
</style>
<!--- End styling code. --->

---
title: "Short Lab 1"
author: "INSERT YOUR NAME HERE"
date: "9/30/2020"
output: html_document
---

To begin, please follow the course links to download R, RStudio, and R Markdown.
You should edit this .Rmd using RStudio, then click *Knit* in the menu bar of the Source window (above the text of this .Rmd).
Remember, you must submit *both* your .Rmd and the compiled .html in order to receive full credit!

## Part 1. Written Response

**1.** What is the name you want me to call you?

Replace this text with your answer.

**2.** What are your gender pronouns? (Example: mine are he/him/his)

Replace this text with your answer.

**3.** What is your major?

Replace this text with your answer.

**4.** What is your coding/computational experience? Please specifically mention whether or not you have any experience in each of R, Rmarkdown, and LaTeX.

Replace this text with your answer.

**5.** In 3 sentences or less, please explain what you hope to take away from this course.

Replace this text with your answer.

**6.** Explain the grading scale of Short Labs in this course.

Replace this text with your answer.

**7.** What is the penalty if you submit an assignment 32 hours late?

Replace this text with your answer.

**8.** What percent extra credit is possible in this course, and how do you earn it?

Replace this text with your answer.

**9.** What time zone will you be in for the duration of this course?

Replace this text with your answer.

**10.** Do my current proposed office hour times work for you? If not, please provide some general availability.

Replace this text with your answer.

<!--- Note: I use the backslash to skip a line for formatting. --->
\ 

## Part 2. R Code

**9a.** Fill in the code below to assign two numeric variables.

```{r}
# Your code here
```

**9b.** Use object type functions to confirm your variables are numeric.

```{r}
# Your code here
```

**10.** Write one line of code that involves an arithmetic operator and both of your variables.

```{r}
# Your code here
```

**11.** Write one line of code that involves a comparison operator, both of your variables, and evaluates to `TRUE`.

```{r}
# Your code here
```

**12.** Write one line of code that involves a comparison operator, both of your variables, and evaluates to `FALSE`.

```{r}
# Your code here
```

**13.** Write one line of code that involves a comparison operator, a logical operator, both of your variables, and evaluates to `TRUE`. Add a comment to explain what your code is doing in words.

```{r}
# Your code here
```

**14.** Write one line of code that involves a comparison operator, a logical operator, both of your variables, and evaluates to `FALSE`. Make each of the operators different than they were in question 13. Add a comment to explain what your code is doing in words.

```{r}
# Your code here
```

**15.** Remove one of your variables from your workspace, then print out your workspace to confirm.

```{r}
# Your code here
```
