---
title: "Short Lab 1"
author: "INSERT YOUR NAME HERE"
date: "1/6/2020"
output: html_document
---


<!--- Please do not touch the code below! It is for styling. --->
<style type="text/css">
/* Whole document: */
body{
  font-family: "Palatino Linotype", "Book Antiqua", Palatino, serif;
  font-size: 12pt;
}
h1.title {
  font-size: 38px;
  text-align: center;
}
h4.author {
  font-size: 18px;
  text-align: center;
}
h4.date {
  font-size: 18px;
  text-align: center;
}
</style>
<!--- Begin here! Fill in text and R code for each of these questions. --->

To begin, please follow the course links to download R, RStudio, and Rmarkdown.
You should edit this .Rmd using RStudio, then click *Knit* in the menu bar of the Source window (above the text of this .Rmd).
Remember, you must submit *both* your .Rmd and the compiled .html in order to receive full credit!

### **Part A. Written Response**

#### *1. What is the name you want me to call you?*

Replace this text with your answer.

#### *2. What are your gender pronouns? (Example: mine are he/him/his)*

Replace this text with your answer.

#### *3. What is your coding/computational experience? Please specifically mention whether or not you have any experience in each of R, Rmarkdown, and LaTeX.*

Replace this text with your answer.

#### *4. In 3 sentences or less, please explain what you hope to take away from this course.*

Replace this text with your answer.

#### *5. Explain the grading scale of Short Labs in this course.*

Replace this text with your answer.

#### *6. What is the penalty if you submit an assignment 32 hours late?*

Replace this text with your answer.

#### *7. What percent extra credit is possible in this course, and how do you earn it?*

Replace this text with your answer.

<!--- Note: I use the backslash to skip a line for formatting. --->
\ 

### **Part B. R Code**

#### *8a. Fill in the code below to assign two numeric variables.*

```{r}
# Your code here
```

#### *8b. Use object type functions to confirm your variables are numeric.*

```{r}
# Your code here
```

#### *9. Write one line of code that involves an arithmatic operator and both of your variables*

```{r}
# Your code here
```

#### *10. Write one line of code that involves a comparison operator, both of your variables, and evaluates to TRUE*

```{r}
# Your code here
```

#### *11. Write one line of code that involves a comparison operator, both of your variables, and evaluates to FALSE*

```{r}
# Your code here
```

#### *12. Write one line of code that involves a comparison operator, a logical operator, both of your variables, and evaluates to TRUE. Add a comment to explain what your code is doing in words.*

```{r}
# Your code here
```

#### *13. Write one line of code that involves a comparison operator, a logical operator, both of your variables, and evaluates to FALSE. Make each of the operators different than they were in question 12. Add a comment to explain what your code is doing in words.*

```{r}
# Your code here
```

#### *14. Remove one of your variables from your workspace, then print out your workspace to confirm.*

```{r}
# Your code here
```
